---
title: "hw2"
output: github_document
---


```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset

```{r}
trashwheel_df = 
  read_xlsx("./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
            sheet = "Mr. Trash Wheel",
            range = cell_cols("A:N")
            ) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls=round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

read precipitation data

```{r}
precip_2018=
  read_xlsx("./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
            sheet = "2018 Precipitation",
            skip =1
            ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year=2018) %>% 
  select(year,everything())

precip_2017=
  read_xlsx("./Trash-Wheel-Collection-Totals-8-6-19.xlsx",
            sheet = "2017 Precipitation",
            skip =1
            ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year=2017) %>% 
  select(year,everything())

```

Combine annual precipitation

```{r}

month_df = tibble(
  month_num = 1:12,
  month_name = month.name
)

precip_df=
  bind_rows(precip_2018,precip_2017)

precip_df = left_join(precip_df,month_df,by = c("month" = "month_num"))
```


Dataset **"trashwheel_df"** contains the information of statistics of each dumpter with the year range of `r range(pull(trashwheel_df,year))`. There are totally `r nrow(trashwheel_df)` of observations. It also contains the weight, volume as well as the number and type of trash collected. 
Dataset **"precip_data"** contains the information of total precipitation in each month with the year range of `r range(pull(precip_df,year))`. There are totally `r nrow(precip_df)` observations.

Total precipitation in 2018 is `r filter(precip_df,year==2018) %>% pull(total) %>% sum()`(in).

The median number of sports balls in a dumpster in 2017 is `r filter(trashwheel_df,year==2017) %>% pull(sports_balls) %>% median()`.


## Problem 2

Read & clean data

```{r}
subway = read_csv("./NYC_Transit_Subway_Entrance_And_Exit_Data.csv",col_types = '-ccnnccccccccccccf-f--l---------') %>% 
  janitor::clean_names() %>% 
  rename(ADA_compliance=ada)
  mutate (subway, entry=as.logical(ifelse(pull(subway, entry)=="YES",1, 0))) 
```

This data contains information and statistics from the trash collectors in Baltimore, Maryland. It include the number of different type of trash at a given time. In this datasets, the dimension is `r nrow(subway)`*`r ncol(subway)`. The file was imported with specified variable type assigned with function "col_types". Then the column name was cleaned with "clean_names()". Column name "ada" was renamed into "ADA_compliance" for clarification. Lastly, variable "entry" was changed to logical variable.

**How many distinct stations are there?**

There are `r distinct(subway, line, station_name, .keep_all = TRUE) %>% nrow()` distinct stations

**How many stations are ADA compliant**

There are `r distinct(subway, line, station_name, .keep_all = TRUE) %>% filter(ADA_compliance == TRUE) %>% nrow()` stations that are ADA compliant.


**What proportion of station entrances / exits without vending allow entrance?**

`r filter(subway,vending=="NO"& entry=="YES") %>% nrow()/filter(subway,vending=="NO") %>% nrow()*100` % of station entrances / exits without vending allow entrance.


reformat data

```{r}
subway_distinct = distinct(subway, line, station_name, .keep_all = TRUE) %>%  
  pivot_longer(
               route1:route11,
               names_to = "route_name",
               names_prefix = "route",
               values_to = "route_number"
               ) %>% 
  drop_na(route_number)
```

**How many distinct stations serve the A train?**

`r filter(subway_distinct,route_number == "A") %>% nrow()` distinct stations serve the A train

**Of the stations that serve the A train, how many are ADA compliant?**

Among stations that serve the A train, `r filter(subway_distinct,route_number == "A"& ADA_compliance == TRUE) %>% nrow()` are ADA compliant.

## Problem 3

clean the pols-month data
```{r}
pols_df = read_csv("./fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon,into = c("year","month","day"),remove = TRUE,convert = TRUE)

pols_df = left_join(pols_df,month_df,by = c("month" = "month_num")) %>% 
  select(-c(month,day)) %>% 
  select(year, month = month_name,everything()) %>% 
  mutate (
    president = case_when(
      prez_dem == 1 ~"dem",
      prez_gop == 1 ~ "gop"
    )
  ) %>% 
  select(-c(prez_dem,prez_gop))
```

clean the snp.csv data

```{r}
snp_df = read_csv("./fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names()%>% 
  separate(date,into = c("month","day","year"),remove = TRUE,convert = TRUE)

snp_df = left_join(snp_df,month_df,by = c("month" = "month_num")) %>% 
  select(-c(month,day)) %>% 
  select(year, month = month_name,everything()) %>% 
  arrange(year,month)
```

clean the unemployment data

```{r}
unemployment_df = read_csv("./fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names()%>% 
rename(January=jan, February=feb,March=mar,April=apr,May=may,June=jun,July=jul,August=aug,September=sep,October=oct,November=nov,December=dec) %>% 
  pivot_longer(
    January:December,
    names_to = "month",
    values_to = "unemployment"
  )
```

merge the datasets

```{r}
combined_df = left_join(pols_df,snp_df,by = c("year","month")) %>% 
  left_join(unemployment_df,by = c("year","month"))

```

The dataset **"pols_df"** contains information of national politicians who are democratic or republican. The key variables include whether president was democratic or republican at a given time, as well as number of governers, senators and representativs from the two parties. The dimension of pols dataset is `r nrow(pols_df)`*`r ncol(pols_df)`. The year range is [`r range(pull(pols_df,year))`].

The dataset **"snp_df"** contains the information of the closing value Standard & Poor's stock market index (S&P) at a given time. The specific date was dropped but year and month were kept to merge with other datasets.The dimension of dataset is `r nrow(snp_df)`*`r ncol(snp_df)`. The year range is [`r range(pull(snp_df,year))`].

The dataset **"unemployment_df"** contains the perceptage of unemployment in a given month and year. The dimension of dataset is `r nrow(unemployment_df)`*`r ncol(unemployment_df)`. The year range is [`r range(pull(unemployment_df,year))`].

The **combined dataset** merged from pols_df, snp_df and unemployment_df provide information on politicians at a given time, and economic performance (represented by S&P) and unemployment at that time. The dimension of the resulting data is `r nrow(combined_df)`*`r ncol(combined_df)`. The year range is [`r range(pull(combined_df,year))`]. The politician affiliation of the president was represented by variable "president". The S&P value was represented by variable "closing" and the unemployment rate was represented by variable "unemployment".
